---
title: "ANOVA and Regression"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: 3
fontsize: 12 pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE) #to supress warning messages in all output
```

***

```{r}
library(mosaic)
library(ggplot2)
library(dplyr)
library(psych)
```

#ANOVA

We are going to be using the Big Five Inventory dataset, `bfi`, to demonsrate ANOVA. This dataset it contained in the `psych` package. 

```{r}
data(bfi)
```

Before we get into the ANOVA, we should first create all of our scale scores. 

```{r}
corr.test(select(bfi, A1, A2, A3, A4, A5))
corr.test(select(bfi, C1, C2, C3, C4, C5))
corr.test(select(bfi, E1, E2, E3, E4, E5))
corr.test(select(bfi, N1, N2, N3, N4, N5))
corr.test(select(bfi, O1, O2, O3, O4, O5))

bfi <- bfi %>%
  mutate(A1.r = (min(A1, na.rm = TRUE) + max(A1, na.rm = TRUE)) - A1,
         C4.r = (min(C4, na.rm = TRUE) + max(C4, na.rm = TRUE)) - C4,
         C5.r = (min(C5, na.rm = TRUE) + max(C5, na.rm = TRUE)) - C5,
         E1.r = (min(E1, na.rm = TRUE) + max(E1, na.rm = TRUE)) - E1,
         E2.r = (min(E2, na.rm = TRUE) + max(E2, na.rm = TRUE)) - E2,
         O2.r = (min(O2, na.rm = TRUE) + max(O2, na.rm = TRUE)) - O2,
         O5.r = (min(O5, na.rm = TRUE) + max(O5, na.rm = TRUE)) - O5)

alpha(select(bfi, A1.r, A2, A3, A4, A5))
alpha(select(bfi, C1, C2, C3, C4.r, C5.r))
alpha(select(bfi, E1.r, E2.r, E3, E4, E5))
alpha(select(bfi, N1, N2, N3, N4, N5))
alpha(select(bfi, O1, O2.r, O3, O4, O5.r))

bfi <- bfi %>%
  mutate(agreeable = (A1.r + A2 + A3 + A4 + A5)/5,
         conscient = (C1 + C2 + C3 + C4.r + C5.r)/5,
         extrov = (E1.r + E2.r + E3 + E4 + E5)/5,
         neurot = (N1 + N2 + N3 + N4 + N5)/5,
         openness = (O1 + O2.r + O3 + O4 + O5.r)/5) %>%
  filter(!is.na(education)) 

glimpse(bfi)
```

Let's also take a look at the distributions of our new variables. 

```{r}
ggplot(bfi, aes(x = agreeable)) + geom_density()
ggplot(bfi, aes(x = conscient)) + geom_density()
ggplot(bfi, aes(x = extrov)) + geom_density()
ggplot(bfi, aes(x = neurot)) + geom_density()
ggplot(bfi, aes(x = openness)) + geom_density()
```

Yesterday we used `t.test()` to test for differences in conscientiousness between those who graduated college, and those who did not. Note that I did not create the variable again, I used a logical statement directly in the `t.test()` function. 

```{r}
t.test(conscient ~ (education > 3), data = bfi)
```

A better thing to do would be to check for differences with a one-way anova.

##One-Way ANOVA

Perform a one-way ANOVA for education level on concientousness. Let's first look at the distributions of conscientiousness by education level. Note that we're piping a datset right into `ggplot()`.

```{r}
bfi %>%
  mutate(education = as.factor(education)) %>%
ggplot(aes(x = education, y = conscient, fill = education)) +
  geom_boxplot(alpha = .5) 
```

Also, let's get descriptives by education level.

```{r}
#descriptives
```

We can perform a Levene's test to test the homogeneity of variance assumption with the `leveneTest()` function that's in the `car` package.

```{r}
#install.packages("car")
library(car)

leveneTest(bfi$conscient, bfi$education)
```

Now that we've checked out assumptions, finally, we can run the one-way ANOVA. 

```{r}
mod1 <- aov(conscient ~ education, data = bfi)

summary(mod1)
```

There are no statistically significant differences between people with different education levels in conscientiousness. We might want a bar graph anyway. 

```{r}
#For nice labels for education level
bfi <- bfi %>%
  mutate(education = as.factor(education),
         education = factor(education, labels=c('HS', 
                                                'finished HS', 
                                                'some college', 
                                                'college graduate', 
                                                'graduate degree')))

#A small companion dataset for making error bars         
plotdata <- bfi %>% 
  group_by(education) %>%
  summarise(mean = mean(conscient, na.rm = TRUE), 
            stdv = sd(conscient, na.rm = TRUE), 
            n = n()) %>%
  mutate(se = stdv/sqrt(n))

#Making the Bar Graph
ggplot(plotdata, aes(x = education, 
                     y = mean, 
                     fill = education)) +
  geom_bar(stat = "identity") +
  geom_errorbar(aes(ymax = mean + se, ymin = mean - se), 
                position = position_dodge(0.9), width = 0.25) +
  labs(x = "Education Level", y = "Conscientiousness") + 
  ylim(0, 5) +  #the scale is really from 1 to 6
  ggtitle("Concientiousness by Education Level") + 
  scale_fill_discrete(guide = FALSE) +
  theme_minimal()
```


##Two-Way ANOVA

First let's do some data stuff we'll need.

```{r}
#install.packages("tidyr")
library(tidyr)

?unite

bfi <- bfi %>%
  mutate(gender = as.character(factor(gender, labels=c('Men', 'Women'))),
         education = as.character(education))
```


Two-way ANOVA for gender by education level on conscientiousness. We can get `favstats()` split by another categorical variable with the `|` symbol. It's above your `return` key.

```{r}
favstats(conscient ~ gender|education, data = bfi)
```

Alternatively we can use `dplyr`

```{r}
bfi %>%
  group_by(gender, education) %>%
  summarise(M = mean(conscient, na.rm = TRUE),
            Md = median(conscient, na.rm = TRUE),
            SD = sd(conscient, na.rm = TRUE))
```

Now we perform the two-way ANOVA.

```{r}

```

Bar graph

```{r}
plotdata <- expdata %>% 
  group_by(bio_sex, gender_expression) %>%
  summarise(mean = mean(hirability), 
            stdv = sd(hirability), 
            n = n()) %>%
  mutate(se = stdv/sqrt(n))

ggplot(plotdata, aes(x = factor(bio_sex), y = mean, fill = factor(gender_expression))) +
  geom_bar(stat = "identity", position = position_dodge(0.9)) +
  geom_errorbar(aes(ymax = mean + se, ymin = mean - se), position = position_dodge(0.9), width = 0.25) +
  labs(x = "bio_sex", y = "hirability_average") +
  ggtitle("hirability by biological sex and gender expression") +
  scale_fill_discrete(name = "gender expression")
```

We will see another example of a two-way ANOVA in the reproducible APA style document.

##Mixed Effects ANOVA

To demonstrate paired samples t-test I'd like to use yet another built in dataset. The `sat.act` dataset has information for 700 people on their SAT verbal, SAT quantitative, and ACT scores.

```{r}
data(sat.act)
```

```{r}
sat.act %>%
  filter(gender == 1) %>%
  t.test(SATV, SATQ, data = ., paired = TRUE)
```

```{r}
sat.act %>%
  filter(gender == 2) %>%
  t.test(SATV, SATQ, data = ., paired = TRUE)
```

Gender by topic two-way mixed effectcs ANOVA

```{r}

```

Gender by education by topic three-way mixed effects ANOVA

```{r}

```

#Regression

What if we wanted to treat `education` like an interval measured variable instead of ordinal? We could then regress concientiousness on education in a simple linear regression model. 

```{r}

```

Let's add gender to make it a multiple regression model. 

```{r}

```

We can also add interactions. In R we actually don't need to specify the main effects, they'll be added for us. difference between `*` and `:`

```{r}

```

##Logistic Regression

Because personality is relatively stable, we may instead ask are people who are more concientious more likely to graduate from college? This would be a logistic regression model. We can use the code we wrote yesterday for creating the dichotomous variable 1 = yes college, 0 = no college. 

```{r}

```



