---
title: "Confirmatory Factor Analysis and SEM"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: 3
fontsize: 12 pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE) #to supress warning messages in all output
```

***

```{r}
library(mosaic)
library(ggplot2)
library(dplyr)
library(psych)
```

```{r}
#install.packages("lavaan")
library(lavaan)

#install.packages("semPlot")
library(semPlot)
```

#Confirmatory Factor Analysis

Now, instead of taking a shot in the dark about the factor structue, we are going in with a clear idea about which items we things load on which factors. Our goal is to test how well our model of the factor structure actually fits the data. 

First we specify the model.

```{r}
bf_model <- ' agreeable =~ A1 + A2 + A3 + A4 + A5
              conscient =~ C1 + C2 + C3 + C4 + C5
              extrov =~ E1 + E2 + E3 + E4 + E5
              neurot =~ N1 + N2 + N3 + N4 + N5
              openness =~ O1 + O2 + O3 + O4 + O5 '
```

##Diagram

Plot the model to make sure it is what you want. Here we are using the `semPaths()` and `semPlotModel()` from the `semPlot` package. The `rotation = 2` option forces the exogenous vairables to be on the left side. 

```{r}
semPaths(semPlotModel(bf_model), rotation = 2)
```

##Output

Then fit the CFA model with `cfa()` and ask for the relevant output.

```{r}
bf_fit <- cfa(bf_model, data = bfi)

summary(bf_fit, fit.measures = TRUE, rsq=TRUE)
inspect(bf_fit, what = "std") #for standardized estimates
```

##Modification Indicies

Because our model is not a great fit to our data, we migth want to ask for the *modification indicies* with the `modindices()` function. `lavaan` will give you a set of changes you can make that will increase the fit on the model. The `mi` value is the amount a reduction in the chi-square statistics that each change would bring---so the higher the better because a reduction in chi-square means and increase in fit! 

```{r}
modindices(bf_fit) %>% 
  arrange(desc(mi)) %>%
  select(lhs, op, rhs, mi) %>%
  head(10)
```

##Re-specifying the Model

Maybe we want to add error covariances based on modificaton indicies. I recommend only doing this if it makes theoretical sense. Also, what ever theoretical rule you use to justify making these error correlations, if it applies to other sets of items, then add the correlations there too. Be consistent! **If you just make changes based on the modification indicies without thinking carefully about the theory behind them, then you are doing exploratory factor analysis.**   

```{r}
bf_model_re1 <- ' agreeable =~ A1 + A2 + A3 + A4 + A5
                  conscient =~ C1 + C2 + C3 + C4 + C5
                  extrov =~ E1 + E2 + E3 + E4 + E5
                  neurot =~ N1 + N2 + N3 + N4 + N5
                  openness =~ O1 + O2 + O3 + O4 + O5 
                    
                  N1 ~~ N2
                  N3 ~~ N4'
```

How would you ask for a diagram of this model?

```{r}
#diagram here.
```

```{r}
bf_fit_re1 <- cfa(bf_model_re1, data = bfi)

summary(bf_fit_re1, fit.measures = TRUE, rsq=TRUE)
inspect(bf_fit_re1, what = "std")
```

```{r}
modindices(bf_fit_re1) %>% 
  arrange(desc(mi)) %>%
  select(lhs, op, rhs, mi) %>%
  head(10)
```

We might also want to remove A1 -- "Am indifferent to the feelings of others," and O4 -- "Spend time reflecting on things," based on crappy loadings.

```{r}
bf_model_re2 <- ' agreeable =~ A2 + A3 + A4 + A5
                  conscient =~ C1 + C2 + C3 + C4 + C5
                  extrov =~ E1 + E2 + E3 + E4 + E5
                  neurot =~ N1 + N2 + N3 + N4 + N5
                  openness =~ O1 + O2 + O3 + O5 
                    
                  N1 ~~ N2
                  N3 ~~ N4'

bf_fit_re2 <- cfa(bf_model_re2, data = bfi)

summary(bf_fit_re2, fit.measures = TRUE, rsq=TRUE)
inspect(bf_fit_re2, what = "std")
```

It is still not a great fit to the data, but I don't think I'd want to amke any of these modifications. 

```{r}
modindices(bf_fit_re2) %>% 
  arrange(desc(mi)) %>%
  select(lhs, op, rhs, mi) %>%
  head(10)
```

We can compare teh CFIs for the 3 models with the `fitmeasures()` function. 

```{r}
fitmeasures(bf_fit)[9]
fitmeasures(bf_fit_re1)[9]
fitmeasures(bf_fit_re2)[9]
```

#Structural Equation Modeling

In real life you should not proceed to the structural equation modeling phase unless you have a good fitting measurement model (CFA). For demonstration purposes let's add gender, age, and education into the model as predictors of the 5 personality factors. 

```{r}
bf_model_sem <- ' agreeable =~ A2 + A3 + A4 + A5
                  conscient =~ C1 + C2 + C3 + C4 + C5
                  extrov =~ E1 + E2 + E3 + E4 + E5
                  neurot =~ N1 + N2 + N3 + N4 + N5
                  openness =~ O1 + O2 + O3 + O5 
                    
                  N1 ~~ N2
                  N3 ~~ N4

                  agreeable ~ age + gender + education
                  conscient ~ age + gender + education
                  extrov ~ age + gender + education
                  neurot ~ age + gender + education
                  openness ~ age + gender + education
            
                  age ~~ gender
                  age ~~ education
                  gender ~~ education
                  '
```

Let's make the graph.

```{r}
semPaths(semPlotModel(bf_model_sem), rotation = 2)
```

Instead of the `cfa()` function we will use the `sem()` function.

```{r}
bf_fit_sem <- sem(bf_model_sem, data = bfi)

summary(bf_fit_sem, fit.measures = TRUE, rsq=TRUE)

#standardized item loadings
inspect(bf_fit_sem, what = "std")$lambda 

#standardized path estimates
inspect(bf_fit_sem, what = "std")$beta 

#correlations bewteen latent variables and exogenous variables
inspect(bf_fit_sem, what = "std")$psi 
```


