---
title: "Exploratory Factor Analysis"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: 3
fontsize: 12 pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE) #to supress warning messages in all output
```

***

```{r}
library(mosaic)
library(ggplot2)
library(dplyr)
library(psych)
```

#Exploratory Factor Analysis

```{r}
data(bfi)
```

```{r}
#correlation matrix of ALL items
r_matrix <- corr.test(select(bfi, -gender, -education, -age))$r
cor.plot(r_matrix)
```

Test for gender differences in concientiouness between those who graduated college, and those who did not. Hint: use `mutate()` to create a dichotomous variable as we did for `wise_hus`. 

```{r}
#more t-Tests
```

A better thing to do would be to check for differences with a one-way anova.

##One-Way ANOVA

Perform a one-way ANOVA for education level on concientousness.

```{r}

```

Bar graph

```{r}

```


##Two-Way ANOVA

Two-way ANOVA for gender by education level on concientiouness

```{r}

```

Bar graph

```{r}

```

We will see another example of a two-way ANOVA in the reproducible APA style document.

##Mixed Effects ANOVA

To demonstrate paired samples t-test I'd like to use yet another built in dataset. The `sat.act` dataset has information for 700 people on their SAT verbal, SAT quantitative, and ACT scores.

```{r}
data(sat.act)
```

```{r}
sat.act %>%
  filter(gender == 1) %>%
  t.test(SATV, SATQ, data = ., paired = TRUE)
```

```{r}
sat.act %>%
  filter(gender == 2) %>%
  t.test(SATV, SATQ, data = ., paired = TRUE)
```

Gender by topic two-way mixed effectcs ANOVA

```{r}

```

Gender by education by topic three-way mixed effects ANOVA

```{r}

```

#Regression

What if we wanted to treat `education` like an interval measured variable instead of ordinal? We could then regress concientiousness on education in a simple linear regression model. 

```{r}

```

Let's add gender to make it a multiple regression model. 

```{r}

```

We can also add interactions. In R we actually don't need to specify the main effects, they'll be added for us. difference between `*` and `:`

```{r}

```

##Logistic Regression

Because personality is relatively stable, we may instead ask are people who are more concientious more likely to graduate from college? This would be a logistic regression model. We can use the code we wrote yesterday for creating the dichotomous variable 1 = yes college, 0 = no college. 

```{r}

```



