---
title: "Correlation Matrices, Reliability Tests, and t-Tests"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: 3
fontsize: 12 pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE) #to supress warning messages in all output
```

***

```{r}
library(mosaic)
library(ggplot2)
library(dplyr)

#install.packages("psych")
library(psych)
```

#Data inside R packages

To demonstrate the use of the `alpha()` fuction in the `psych` package from computing Cronbach's alpha, we will need a data set with scale items. Many packages have example datasets pre-loaded. The `psych` package has a dataset called `bfi` that has n = 2800 participants responding to 25 personality itmes for the Big Five Inventory. 

```{r}
#We can see a list of ALL the datasets current avaliable
data()
```

If we see a dataset we like we can load it into our environment 

```{r}
data(bfi)
#?bfi
glimpse(bfi)
```

#Correlation Matrix

First, we can get descriptive statistics for individual personality scale items. Let's do the agreeableness scale first. 

```{r}
#Get descriptive statistics with mosaic or dplyr here. Get them for Age too.

#Also, get the frequencies for gender and education 

```

In addition to descriptive statistics, we would probably also like to get a correlation matrix for items on a scale. I like the function `corr.test()` in the `psych` package. Notice that we are using the `dplyr` `select()` function inside of `corr.test()`.

```{r}
corr.test(select(bfi, A1, A2, A3, A4, A5))
```

There is also a `pairs()` fucntion in Base R and the `ggpairs()` function in the `GGally` package that is a ggplot correlation matrix/matrix scatterplot provider.

```{r}
#install.packages("GGally")
library(GGally)

ggpairs(select(bfi, A1, A2, A3, A4, A5))
```

It is clear that we will need to reverse score A1 to assess reliability and to create our agreeableness scale scores. Hint: to reverse score an item use this formula: `A1.r = (min(A1, na.rm = TRUE) + max(A1, na.rm = TRUE)) - A1`.

```{r, include=FALSE}
#Use a dplyr funcion to add a new variable A1.r, the reversed item. 
```

Now compute correlation matrices for the other 4 personality subscales and create any reverse scored items the need to be created.

```{r}
#correlation matrices and reverse scored items
```

For more practice, find another dataset that sounds interesting and run some correlation matricies.

```{r}
data()
```

#Cronbach's alpha

Now that we have our reverse scored items and we have a sense of how the items relate, the next step is to test the scale reliability. We do this with Cronbach's alpha. There is a function called `alpha()` in the `psych` package. 

```{r}
alpha(select(bfi, A1.r, A2, A3, A4, A5))
```

With an alphas of .70 we have a reliabile agreeableness scale. The next step would be to create agreeableness scale scores for participants, the average of these 5 items. We can do this with `mutate()`. Note that we cannot use `mean()` here because `mean()` is going to get the mean of a column (a variable). What we want it the mean across columns.   

```{r}
bfi <- bfi %>%
  mutate(agreeable = (A1.r + A2 + A3 + A4 + A5)/5)
```

Let's take a look at it with a visualization. Also try facetting by gender.

```{r}
#make some histograms, boxplots, and density plots of agreeableness
```

Now calculate the reliabilities for each of the other 4 scales. Be sure to include the reverse scored items you created where relevant. 

```{r}
#calculate reliabilities
```

Create scale scores and make visualizations of those scale scores. 

```{r}
#scale scores and visualizations
```


#t-Tests

##Independent Samples t-Test

Next we can ask, are men and women different on agreeableness? To answer this question we need to run an independent samples t-test. We'll use the function `t.test()` in the `mosaic` package. Recall that `mosaic` first needs the formula, `y ~ x`, then the data, `data = dataName`.

```{r}
t.test(agreeable ~ gender, data = bfi)
```

There is a statistically significant difference between women and men on agreeableness, *t*(1654.50) = -10.73, *p* < .001 , with women (*M* = 4.77, *SD* = 0.86) scoring higher than men (*M* = 4.38, *SD* = 0.93). It is possible to code in these numbers such that if the data were updated, the text would update as well. 

```{r}
tmod <- t.test(agreeable ~ gender, data = bfi)
ds <- favstats(agreeable ~ gender, data = bfi)
```

There is a statistically significant difference between women and men on agreeableness, *t*(`r round(tmod$parameter, 2)`) = `r round(tmod$statistic, 2)`, *p* < `r ifelse(tmod$p.value > .001, round(tmod$p.value, 3),.001)`, with women (*M* = `r round(ds[2,7], 2)`, *SD* = `r round(ds[2,8], 2)`) scoring higher than men (*M* = `r round(ds[1,7], 2)`, *SD* = `r round(ds[1,8], 2)`).

We can also ask for equal variance assumed. 

```{r}
t.test(agreeable ~ gender, data = bfi, var.equal = TRUE)
```

Test for gender differences for the other 4 scales. Then test differences between those who graduated college, and those who did not. Hint: use `mutate()` to create a dichotomous variable as we did for `wise_hus`. 

```{r}
#more t-Tests
```


##Paired Samples t-Test

To demonstrate paired samples t-test I'd like to use yet another built in dataset. The `sat.act` dataset has information for 700 people on their SAT verbal, SAT quantitative, and ACT scores.

```{r}
data(sat.act)
#?sat.act
glimpse(sat.act)
```

Let's compare poeple's scores on the SAT verbal to their scores on the SAT quantitative using the `t.test()` function. This time instead of using the formula in the beginning, we add the two variables to be compared separated by a `,`. We also need to add `paired = TRUE`.

```{r}
t.test(SATV, SATQ, data = sat.act, paired = TRUE)
```

There is no staistically significant difference between SAT verbal and SAT quantitative scores, *t*(686) = 0.57, *p* = .566. But perhaps there is a difference for boys only, or for girls only. We can create a dataset of just boys using `filter()` and then another one of just girls. This would be completely fine, but we could also pipe directly into the `t.test()` function. There is a catch: The pipe, `%>%`, puts the resulting dataframe into the first position in the destination function, but we want it in the third position. We can use the `.` symbol to pipe the dataframe into whichever position we'd like. 

```{r}
sat.act %>%
  filter(gender == 1) %>%
  t.test(SATV, SATQ, data = ., paired = TRUE)
```

```{r}
sat.act %>%
  filter(gender == 2) %>%
  t.test(SATV, SATQ, data = ., paired = TRUE)
```

Explore the differences in `SATV` and `SATQ` for each level of the `education` variable. 

```{r}
#differences for each level of education
```

