---
title: "Intro to dplyr"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: 3
fontsize: 12 pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE) #to supress warning messages in all output
```

***

```{r}
acitelli <- read.csv(file.choose(), header=TRUE)
```

#Data Manipulation (data cleaning)

We'll use the package `dplyr`. The `dplyr` package contains the functions for all of the data cleaning verbs: `filter()`, `mutate()`, `rename()`, `arrange()`, `select()`, `summarize()`, and `group_by()`. You can find a cheat sheet for `dplyr` [here](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf).

```{r, message=FALSE, warning=FALSE}
#install.packages("dplyr")
library(dplyr)
```

##Filtering cases with filter()

First, let's filter cases. We can make a dataset of men only. Notice that we used a double equal sign, `==`, instead of single, `=`. When you want to ask if something is equal to some value or another variable, that is, you want to use equal in a *logical statement*, you need the double equal. You can also use `>` `<` `>=` `<=` `&`, which means **AND**, and finally, `|` which means **OR**.

```{r}
menOnly <- filter(acitelli, gender == 1)
```

We could also use the pipe, `%>%`. 

```{r}
menOnly <- acitelli %>%
  filter(gender ==1)
```

We can save this new data set in our files as a csv. Note that the file path conventions are different for Windows machines. If copying and pasting from properties, you will need to change backslashes to forward slashes. 

```{r}
write.csv(menOnly, 
  file = "/Users/randigarcia/Desktop/Data/men.csv") #replace file path
```

It's in your working directory, which you can find with the following command. It is an empty function, so it looks weird.

```{r, eval=FALSE}
getwd()
```

How about only the men who are above the median for `Yearsmar`. First, find the median years married.

```{r}
#use a function in the mosaic package
```

Then, filter for men above that cut off point.

```{r}
mature_hus <- menOnly %>% 
  filter(Yearsmar > -1.089)
```

Instead of first finding the median with `favstats()`, we could ask for the median inside of `filter()` with the Base R function, `median()`. Base R has all of the descriptive stats functions you'd expect, `mean()`, `sd()`, `cor()`, but be careful because if you have missing data you'll have to add `rm.na = TRUE` as an argument to the function. The syntax also differs from `mosaic`.

```{r}
mature_hus <- menOnly %>% 
  filter(Yearsmar > median(Yearsmar))
```

##Adding new variables with `mutate()` 

Let's add a new categorical variable that marks the median split on Yearsmar. After you create it, take a look at it.

```{r}
menOnly <- menOnly %>%
  mutate(mature_hus = Yearsmar > median(Yearsmar))
```

How would you get the frequencies on this variable?

```{r}
#frequencies
```

Now for a sanity check, how would you get the descriptive stats split by mature and non-mature husbands?

```{r}
#descriptives split
```

##Renaming variable with `rename()`

We copy a variable and give it a new name with a function you already know, `mutate()`.

```{r}
menOnly <- menOnly %>%
  mutate(old_hus = mature_hus)
```

We can rename a variable without creating a new one with `rename()`. This is handy if you forget to name variables in Qualtrics!

```{r}
menOnly <- menOnly %>%
  rename(wise_hus = old_hus)
```

We can rename a bunch at the same time. This is handy if you forget to name variables in Qualtrics!

```{r}
menOnly <- menOnly %>%
  rename(self_positivity = self_pos,
         other_positivity = other_pos)
```


##Sorting with `arrange()`

First, you should know that you can sort in the viewer by clicking the (faint) arrows just to the right of each variable name. Give it a try. It's often handy to have a sort command in your code, and/or you might want to sort by more than one variable. 

```{r}
head(acitelli)
```

Say we want to take a peak at the women with the bottom 6 `self_pos` scores. 

```{r}
acitelli %>%
  arrange(gender, self_pos) %>%
  head()
```

We could also save the arranged dataset.

```{r}
acitelli <- acitelli %>%
  arrange(gender, self_pos)

head(acitelli)
```

What about the top 6? We can use the `desc()` function inside of `arrange()`.

```{r}
acitelli %>%
  arrange(gender, desc(self_pos)) %>%
  head()
```


##Selecting variables with `select()`

Save a smaller subset of variables.

```{r}
small <- acitelli %>%
  select(cuplid, gender, satisfaction, self_pos)
```

We can also save everything but some variable(s).

```{r}
no_tension <- acitelli %>%
  select(-tension)
```


##Descriptive statistics with `summarize()`

```{r}
acitelli %>%
  summarize(mean = mean(satisfaction),
            sd = sd(satisfaction),
            min = min(satisfaction))
```

##Grouped descriptives with `group_by()`

We can split the file and view results grouped by some variable.

```{r}
acitelli %>%
  group_by(gender) %>%
  summarize(mean = mean(satisfaction),
            sd = sd(satisfaction),
            min = min(satisfaction))
```

You can use `group_by()` to create aggregated variables, this is handy if you have nested data. We actually do have married couples here, so let's create a dyad mean tension variable. 

```{r}
acitelli <- acitelli %>%
  group_by(cuplid) %>%
  mutate(tension_mean = mean(tension)) %>%
  ungroup()

#this last command is not entirely nessesary, but good practice
```


##Pipelines

We now seen our first pipelines, using `group_by()`. Now we can make a pipeline of many of the commands I did above. The last thing I do is drop useless `gender` variable, because the resulting dataset if all men.

```{r}
mature_hus2 <- acitelli %>%
  filter(gender == 1) %>%
  mutate(wise_hus = Yearsmar > median(Yearsmar)) %>%
  rename(self_positivity = self_pos,
         other_positivity = other_pos,
         personID = cuplid) %>%
  arrange(wise_hus) %>%
  select(-gender)
```

Save a dataset of women who are perceiving above the mean tension, and drop the `simhob` variable.

```{r}
#above the mean
```

What are the couple ID's of the couples with the lowest 3 average satisfaction scores?

```{r}
#3 lowest
```


