##Visualizing Data

Let's make a histogram. We'll use the package `ggplot2` to make our visualizations.

```{r, fig.width = 6, fig.height = 4}
qplot(x = Sat, data = riggsi, bins = 20)
```

Scatterplot. `qplot`, which stands for quick plot, guesses which kind of figure you want. 

```{r, fig.width = 6, fig.height = 4}
qplot(x = Avoid, y = Sat, data = riggsi)
```

Side-by-side boxplots. In this case it does NOT know what to do, so we tell it we want boxplots. It will also be helpful to have a gender variable that is stored as categorical string variable. We do that first.

```{r, fig.width = 6, fig.height = 4}
qplot(y = Sat, x = Genderstring, data = riggsi, geom = "boxplot")
```

We can add a regression line.

```{r, fig.width = 6, fig.height = 4}
qplot(x = Avoid, y = Sat, data = riggsi) + geom_smooth(method = "lm", se = 0)
```

We can also split by gender. Note that for more complex models we need to move away from using the `qplot` function in favor of the heavy duty `ggplot` function. 

```{r, fig.width = 6, fig.height = 4}
ggplot(riggsi, aes(x = Avoid, y = Sat, group = Genderstring, color = Genderstring)) +
         geom_point() + 
         geom_smooth(method = "lm", se = 0)
```
 
You can find a cheat sheet for `ggplot2` here: https://www.rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf.  
 
##Statistical Modeling and Inference in R

Independent samples t-test.

```{r}
t.test(Sat ~ legal, data = menOnly, var.equal = TRUE)
```

Paired samples t-test.

```{r}
t.test(Sat, Avoid, data = menOnly, paired = TRUE)
```

Linear regression.

```{r}
mod <- lm(Sat ~ Avoid + legal, data = menOnly)

summary(mod)
```

Regression diagnostics.

```{r, fig.width = 6, fig.height = 4}
plot(mod)
```

