```{r}
a_thing <- 4
another_thing <- 1
another_Thing <- 7

both_things <- a_thing + another_thing
```

Then we created a tiny data set.

```{r}
a_data_thing <- data.frame(x = 2, y = 8)

a_data_thing$x

#How would we print the variable y? Type your answer below on line 17.


```

Write notes for yourself in the white space. Maybe explain to your future self what dollar signs do.


**ERASE THIS AND TYPE HERE**


Enough playing around, let's load some data!

```{r}
riggsi <- read.csv(file.choose(), header=TRUE)
```

Next, you want to look at your data.

```{r}
head(riggsi)

str(riggsi)

names(riggsi)
```

There is also documentation about functions.

```{r, eval=FALSE}
?head
```

You probably also want descriptive statistics.

```{r}
summary(riggsi)
```

We can also select pieces of a data frame. That first number is the row, the second is the column.

```{r}
riggsi[2, 6]

#You try it! Find a numder you want to pull from the dataset.
#riggsi[ ?, ?]

```

If it is instead a single variable, you can also select a piece.

```{r}
riggsi$Sat[2]
```


***
#Bonus Intro to R

More descriptive stats and frequencies with sample proportions. We'll use the package `mosaic`. You can find more information and a cheat sheet for `mosaic` at [this website](https://cran.r-project.org/web/packages/mosaic/vignettes/MinimalR.pdf).

```{r, message=FALSE, warning=FALSE}
library(mosaic)
```

```{r}
favstats(~Sat, data = riggsi)

tally(~Gender, data = riggsi, format = "proportion")
```

Descriptives split by gender.

```{r}
favstats(Sat ~ Gender, data = riggsi)
```
