---
title: "Intro to R"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
fontsize: 12 pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE) #to supress warning messages in all output
```

#Introduction to R

```{r}
a_thing <- 4
another_thing <- 1
another_Thing <- 7

both_things <- a_thing + another_thing
```

Then we created a tiny data set.

```{r}
a_data_thing <- data.frame(x = 2, y = 8)

a_data_thing$x
```

How would we print the variable y? Type your answer in the chunk below

```{r}

```

Write notes for yourself in the white space. Maybe explain to your future self what dollar signs do.


**ERASE THIS AND TYPE SOME NOTES HERE**


Enough playing around, let's load some data!

```{r}
acitelli <- read.csv(file.choose(), header=TRUE)
```

Next, you want to look at your data.

```{r}
head(acitelli)

str(acitelli)

names(acitelli)
```

There is also documentation about functions.

```{r, eval=FALSE}
?head
```

You probably also want descriptive statistics.

```{r}
summary(acitelli)
```

We can also select pieces of a data frame. That first number is the row, the second is the column.

```{r}
acitelli[2, 6]

#You try it! Find a numder you want to pull from the dataset.
#riggsi[ ?, ?]

```

If it is instead a single variable, you can also select a piece.

```{r}
acitelli$satisfaction[2]
```

In the chunk below, pick out the gender of the person in the 50th case.

```{r}
#try it by referring to the row and column of the data frame.

#try it by referring to the variable, using the dollar sign notation.

```

#Installing Packages 

You might want to get descriptive stats or frequencies for specific variables. There are `base R` functions, but I like to use the package `mosaic`. You can find more information and a cheat sheet for `mosaic` at [this website](https://cran.r-project.org/web/packages/mosaic/vignettes/MinimalR.pdf).

First we need to install the `mosaic` package using the `install.packages()` function. The package name goes inside of the paratheses in double quotes: '"mosaic"'.  This is something we do only once in the console, you wouldn't want to save it in your .Rmd file.

```{r}
#install.packages("mosaic")
```

Once a package is installed, any time we start a new R session and we want to use functions inside of that package, we will need to load the package with the `library()` function. 

```{r, message=FALSE, warning=FALSE}
library(mosaic)
```

#Basic Descriptive Statistics with `mosaic`

The function `favstats()` will give descriptive statistics for a numerical vairable, and the function `tally()` will give you frequencies for a categorical variable (or a numberical variable...if you want it). Functions in `mosaic` use the formula syntax, where `y ~ x`, or for a single variable, `~x`. The `~` key can be found just below your `esc` key. The first argument is the formula, and the second argument is the data frame, e.g., `data = acitelli`.

```{r}
favstats(~satisfaction, data = acitelli)

tally(~gender, data = acitelli)

#tally() can also give you percentages
tally(~gender, data = acitelli, format = "percent")
```

Descriptives split by gender.

```{r}
favstats(satisfaction ~ gender, data = acitelli)
```

What are the standard deviations of perceived `tension` by gender?

```{r}

```

What is(are) the mode(s) of the `self_pos` variable?

```{r}

```

The `mosaic` package also has a function for getting the correlation coefficent, it's called `cor()`. Using the same format (i.e., formula then data), how would you get the correlation of satisfaction and tension?

```{r}

```

