---
title: "Intro to dplyr"
output:
  html_document:
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: 3
fontsize: 12 pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE) #to supress warning messages in all output
```

***

```{r}
acitelli <- read.csv("/Users/randigarcia/Desktop/acitelli.csv", header=TRUE)
```

#Visualizing Data

There are quite a few ways to make figures in R, we'll the popular package `ggplot2`. The easiest way to make a figure with `ggplot2` is with the `qplot()` function. This stands for *quick plot*.  

##Quick Histogram

First, let's make a histogram for satisfaction. Notice in the code below that we did not specify histogram, `qplot()` guesses which type of plot we want based on the variable's type. 

```{r, fig.width = 6, fig.height = 4}
qplot(satisfaction, data = acitelli)
```

There are too many bins (it defaults to 30 bins), we can ask for a specific number by adding the `bins = ` argument. Try playing around the bin number below to find the optimal plot. I put `30` in there as a placeholder.

```{r}
qplot(satisfaction, data = acitelli, bins = 30)
```

##Quick Scatterplot

We might also want a scatterplot. Again,`qplot()` guesses what we want, but it's a good idea to specify which variable goes on the x-axis and which goes on the y-axis. 

```{r, fig.width = 6, fig.height = 4}
qplot(x = tension, y = satisfaction, data = acitelli)
```

We can even add a third variable, mapping it to color. To get the behavior we want, gender has to be a categorical variable, to do this we can use the `as.factor()` function.

```{r, fig.width = 6, fig.height = 4}
qplot(x = tension, y = satisfaction, color = as.factor(gender), data = acitelli)
```

##Quick Boxplots

We can ask for side-by-side boxplots when our x variable is categorical. In this case `qplot()` does **NOT** know what to do, so we tell it we want boxplots with `geom = "boxplot"`. 

```{r, fig.width = 6, fig.height = 4}
qplot(y = satisfaction, x = wise_hus, data = menOnly, geom = "boxplot")
```

##`ggplot()`

For more complex figures we will need to move away from using the `qplot` function in favor of the heavy duty `ggplot` function. 

```{r, fig.width = 6, fig.height = 4}
ggplot(riggsi, aes(x = Avoid, y = Sat, group = Genderstring, color = Genderstring)) +
         geom_point() + 
         geom_smooth(method = "lm", se = 0)
```
 
You can find a cheat sheet for `ggplot2` here: https://www.rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf.  
 
##Statistical Modeling and Inference in R

Independent samples t-test.

```{r}
t.test(Sat ~ legal, data = menOnly, var.equal = TRUE)
```

Paired samples t-test.

```{r}
t.test(Sat, Avoid, data = menOnly, paired = TRUE)
```

Linear regression.

```{r}
mod <- lm(Sat ~ Avoid + legal, data = menOnly)

summary(mod)
```

Regression diagnostics.

```{r, fig.width = 6, fig.height = 4}
plot(mod)
```

